---
title: "ReproducibleProject2"
author: "Kumar Shaket"
date: "16/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Analysis on NOAA Storm Database to identify Severe Weather Events

## Introduction

The basic goal of this assignment is to explore the NOAA Storm Database and answer some basic questions about severe weather events. You must use the database to answer the questions below and show the code for your entire analysis. Your analysis can consist of tables, figures, or other summaries. You may use any R package you want to support your analysis.

## Synopsis
This report will provide a better insight into Storms and severe weather events in United States and the Fatalities, Injuries and property damages left behind. Two questions to be answered:

1 - which types of events are most harmful to population health?

2 - which types of events have the greatest economic consequences?

#Data Processing 

## Loading NOAA Data from CSV File
 1. Download data from Coursera website
 2. unzip file and convert it to csv
 3. read csv file 
 ```{r, cache=TRUE}
stormdata <- read.csv2("/Users/kumarshaket/Downloads/repdata-data-StormData.csv",sep = ",",header = TRUE)
```
```{r, cache=TRUE}
names(stormdata)
```

```{r, cache=TRUE}
head(stormdata,5)
```

## Processing data for analysis

## 1. Event Impact on Population Events

1.1 Converting fatalities from factor to numeric
```{r, cache=TRUE}
stormdata$FATALITIES <- as.numeric(stormdata$FATALITIES)
```

1.2 Calculating aggregate of fatalities for each environment
```{r, cache=TRUE}
FatalityperEvType <- aggregate(FATALITIES~EVTYPE,stormdata,sum)
dim(FatalityperEvType)
```

1.3 Ordering Fatalities due to top 10  events
```{r,cache=TRUE}
Fatalities10events <-FatalityperEvType[order(-FatalityperEvType$FATALITIES),][1:10,]
Fatalities10events
```

1.4 Plot for Fatalities due to top 10 events
```{r,echo=TRUE}
par(mfrow = c(1,1), mar = c(12, 4, 3, 2), mgp = c(3, 1, 0), cex = 0.8)
barplot(Fatalities10events$FATALITIES,names.arg = Fatalities10events$EVTYPE,las=3,col=c("#E69F00", "#56B4E9", "#009E73"),ylab = "Number of Fatalities", main = "10 Fatalities Highest Events")
```

## 2. Event impacting Economies

2.1 Converting PropertyExpense from factor to numeric

```{r}
stormdata$PROPDMGEXP=as.numeric(stormdata$PROPDMGEXP)
```

2.2 Calculating Property Expense due to Event

```{r}
PROPDMGEvtype <- aggregate(PROPDMGEXP ~ EVTYPE,stormdata,sum)
dim(PROPDMGEvtype)
```

2.3 Ordering Property Expense due to top 10 event

```{r}
PROPDMG10Events <- PROPDMGEvtype[order(-PROPDMGEvtype$PROPDMGEXP),][1:10,]
PROPDMG10Events
```
2.4 Plot for Property Damage Expense due to top 10 events 

```{r,echo=TRUE}
par(mfrow = c(1,1), mar = c(12, 4, 3, 2), mgp = c(3, 1, 0), cex = 0.8)
barplot(PROPDMG10Events$PROPDMGEXP,names.arg = PROPDMG10Events$EVTYPE,las=3,col=c("#E69F00", "#56B4E9", "#009E73"),ylab = "Property Damage Expense", main = "10 Property Highest Expense due to Event")
```

## 3. Crop Damage
3.1 Converting CropDamage from factor to numeric
```{r}
stormdata$CROPDMGEXP=as.numeric(stormdata$CROPDMGEXP)
```

3.2 Calculating Crop Damage due to Event
```{r}
CROPDMGEvtype <- aggregate(CROPDMGEXP ~ EVTYPE,stormdata,sum)
dim(CROPDMGEvtype)
```

3.3 Ordering Crop Damage due to top 10 event

```{r}
CROPDMG10Events <- CROPDMGEvtype[order(-CROPDMGEvtype$CROPDMGEXP),][1:10,]
PROPDMG10Events
```
2.4 Plot for Crop Dmaage Expense due to top 10 events 

```{r,echo=TRUE}
par(mfrow = c(1,1), mar = c(12, 4, 3, 2), mgp = c(3, 1, 0), cex = 0.8)
barplot(CROPDMG10Events$CROPDMGEXP,names.arg = CROPDMG10Events$EVTYPE,las=3,col=c("#E69F00", "#56B4E9", "#009E73"),ylab = "Crop Damage Expense", main = "10 Highest Crop Damage due to Event")
```

##Results
1.1 : Across the United States, which types of events (as indicated in the 
EVTYPE(EVTYPE variable) are most harmful with respect to population health?
1.2 : Across the United States, which types of events have the greatest economic consequences?

Ans : **HAIL**